Insurance choice, decision assistance, and insurer steering

### 2021-03-25 finals/lunch-and-learn/lunch-learn-202103.Rmd
update lunch and learn presentation

---
title: "Insurance choice, decision assistance, and insurer steering"
subtitle: "Ian McCarthy & Evan Saltzman"
author: Emory University, Lunch and Learn
date: March 30, 2021 #"`r format(Sys.time(), '%d %B %Y')`"
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
output:
#  html_document: default (toggle on for "simplified" view)
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js","https://platform.twitter.com/widgets.js"]
#knit: pagedown::chrome_print
---

@@

.center[
  ![](https://twitter.com/DaveEvansPhD/status/1234506064470716416?s=20)
]

@@

>>>>

---
title: "Insurance choice, decision assistance, and insurer steering"
subtitle: "Ian McCarthy & Evan Saltzman"
author: Emory University, Lunch and Learn
date: March 29, 2021 #"`r format(Sys.time(), '%d %B %Y')`"
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
output:
#  html_document: default (toggle on for "simplified" view)
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js","https://platform.twitter.com/widgets.js"]
#knit: pagedown::chrome_print
---

@@

<center>
<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr"><a href="https://twitter.com/DaveEvansPhD/status/1234506064470716416">March 2, 2020</a></blockquote><script async src="//platform.twitter.com/widgets.js" charset="utf-8"></script>
</center>

# What is a dominated choice?

In our context...

- Subsidies available for households with incomes below 250% of the FPL
- At 150% of the FPL, larger subsidies are such that actuarial value of silver plans exceeds that of gold and platinum plans
- Gold and platinum plans dominated by silver plans for incomes $\leq$ 150% FPL

About 3.2% of households make a dominated choice in our data

---
# Analysis of dominated choices

.center[
  ![](../../figures/dominated_choice_regression.png)
]

@@

---
# Estimand of interest

- Want to identify the Average Treatment Effect on the Treated (ATT) of decision assistance

- Denote by $Pr[Y_{ij}^{1}]$ the probability of household $i$ selecting plan $j$ with decision assistance

- Denote by $Pr[Y_{ij}^{0}]$ the probability of household $i$ selecting plan $j$ without decision assistance

$$ATT = Pr[Y_{ij}^{1} | \text{Assist}] - Pr[Y_{ij}^{0} | \text{Assist}]$$

---
# Estimation

1. Estimate nested logit model among those without any decision assistance, $\hat{\theta}$

2. Use $\hat{\theta}$ to form predicted counterfactual, $\widehat{Pr[Y_{ij}^{0} | \text{Assist}]}$

3. Use predicted value in ATT estimate: $$\underbrace{Pr[Y_{ij}^{1} | \text{Assist}]}_{Observed} - \underbrace{\widehat{Pr[Y_{ij}^{0} | \text{Assist}]}}_{Predicted}$$

4. Aggregate to level of insurer and metal

# Estimation

- Focus only on new enrollees to avoid presence of inertia

- Estimate separately by year and rating area

---
# Inference

- Clustered bootstrap, by year and rating area

- 200 bootstrap replications

- Confidence interval based on the 5th and 95th percentile of bootstrap replications

---
# Effects of decision assistance

.pull-left[
  ![](../../figures/choice_metals.png)
]

.pull-right[
  ![](../../figures/choice_insurer.png)
]
